Asymmetric relationships and their effects on coexistence: Supporting information

### 2023-07-19 paper/RefereeResponses1.tex

\noindent ***\emph{We thank the referee for this viewpoint. We also thank the referee for acknowledging that this
is their opinion, and that data are insufficient for a clear conclusion to be reached at this point. We find that 
such an acknowledgment is both rare from a referee and very useful for causing us, as authors, to most usefully revise our 
paper. We have added new discussion material covering these topics and find that the new material improves the paper.
We tried to represent both viewpoints in the new material, and we cited the papers the referee mentions.}

\emph{We point out that we worked not only
with the lottery model, but also with an empirical diatom system. That system definitely supported the importance of 
ATAs for coexistance, which would tend to give at least one data point in favor of the importance of these phenomena, though we 
acknowledge that our lab systems is also a system ``where niche differentiation via ... predators is impossible.''
It may also be a system ``where niche differentiation via different resources ... is impossible,'' though resource 
differentiation can happen when you don't expect it. It seems to us 
the first step in assessing the importance of a newly recognized phenomenon has to be examining it in simple systems
such as we have used, prior to examing it in field systems, in all their complexity, where niche differentiation
by resources and predators can more readily occur. But once our paper is published we hope and expect that we and
others will more broadly assess the potential importance of ATAs.}

\emph{We also point
out (and the referee already seems to accept this) that it is possible, \emph{a priori}, for ATAs to be 
important for coexistence in a system even while storage effects are not.
Storage effects are ATA effects plus what we called, in the manuscript, the effects of $EC$ correlation \emph{per se} 
($\Delta_i^{[E \nat C]}$ in our notation). Either of these two summands can be negative or positive.
So if the two effects are opposite in sign and similar in magnitude, storage effects will be unimportant for 
coexistence, but ATA effects will be important.}

\emph{So the referee appears to agree with us that storage effects and ATA effects may or may not turn
out to be broadly important in ecology. Our research goals were to introduce the idea of ATA effects on
coexistence and to show that ATAs \textbf{may} be broadly important for coexistence, by showing examples where they
\textbf{are} important. The few examples we showed are all that are readily available at this time.
We won't know for sure if ATAs (or storage effects) \textbf{are} important until more data can be brought to bear. 
Future work can assess, broadly, how important ATAs are by looking across a larger number of systems, once
available. These are two different goals. It seems crucial to us not to allow insufficiently tested 
opinions about the second set of goals to obstruct the publication of results pertinent to the first set of 
goals, so we thank the referee for not doing that. We modified a few key points in the text to clarify 
our viewpoint that we do not think our work definitively indicates that ATAs will certainly be a crucial
mechanism of coexistence; instead we just think our work opens that possibility, for future exploration.
For instance, we replaced the last sentence of the first paragraph of the Discussion, which used to be
``Though future work should seek to understand precisely when ATAs are or are not important
for coexistence, our work demonstrates the overall importance of this new mechanism'' with the revision
``Future work should seek to understand precisely when ATAs are or are not important
for coexistence. Our work demonstrates the potential for ATAs to be an important mechanism of coexistence, generally.''
And we changed the last sentence of the abstract, which used to read ``ATA influences are an important new 
mechanism of coexistence'' so that it now reads ``ATA influences may be an important new mechanism of coexistence.''
We think these changes make the paper stronger, so thanks for spurring us to make them. The last
sentence of the Introduction already read ``Overall, our study presents a new mechanism of species coexistence
and a means of understanding its theoretical and empirical importance,'' which accurately reflects our belief
that what we've accomplished is an indication that ATAs \textbf{may} be important for coexistence, 
and a means of assessing
that in future work, rather than a definitive conclusion that they are important for coexistence.}

\emph{The new Discussion material is in an SI section which is cited from the main text Discussion section. 
We would have liked to put some of the new material into the main text. But length limits, together with
clear requests from other referees to add a variety of content to the main text, made doing so impossible.} \\
%DONE

@@

\noindent C16) L41: The “we” is slightly confusing because the author teams share only one author \\

\noindent ***\emph{We changed the wording here.} \\
%DONE

\noindent C17) L48: I do not know these on the top of my head, and likely many other readers will neither, so maybe quickly refresh our memory \\

\noindent ***\emph{There is no space to intro these things, so we limited this text to just mentioning 
extinction risks and stability of ecosystem functioning. Whereas some ecologists may not have heard of 
Taylor's law, we feel confident that extinction risks and stability of ecosystem functioning will ring
a bell for most ecologists. It was not actually necessary to link to Taylor's law, we just though it 
would provide an additional tie-in to another part of the literature. But that won't work if the referee is right 
that most ecologists will become confused, so we were happy to drop that connection here.} \\
%DONE

@@

\noindent C18) L49: This is not immediately clear to me why (obviously, you’ve shown this in the other paper). Is it possible to briefly state why? \\

\noindent ***\emph{It is not necessary for the reader to understand the mechanism we explored in that other
paper in order to understand this current paper, we were just trying to tie the current work in to 
another area of research in ecology, for the sake of contextualizing it a bit and motivating studies of ATAs generally. 
So an explanation, here, is not 
strictly necessary to the current paper. Nevertheless, we tried to satisfy the referee's request of adding a bit of explanation
without disrupting the flow too much. Thanks for your interest, by the way, we encourage you to have a look at the
other paper if you haven't already.} \\
%DONE

@@

\noindent C19) L61: What if not? Or many readers have only rough understanding of it, maybe quickly refresh or refer to later \\

\noindent ***\emph{It is not necessary for the reader to understand what these are at this point, and,
in fact, the reader does not need to understand what relative nonlinearity is at all for the main text
and main ideas of the paper. We apologize for giving the false impression, here, that more background knowledge was needed
than is actually the case. We eliminated the reference to relative nonlinearity and indicated to the 
reader that storage effects would be conceptually reviewed below (a couple paragraphs down) and then 
defined in Theory, which they are.} \\
%DONE

\noindent C20) Sidenote: After writing these few comments it appears to me that you seem to have a very knowledgeable reader in mind, which may not be the case for a general journal as Ecology letters. \\

\noindent ***\emph{Well, we appreciate the referee's effort to spur us to produce the clearest 
paper possible, and we have used the comments to clarify. But, of the comments 
C16-C19 above, only C19 relates at all to whether this paper is written for a general ecological audience,
and we have revised the text C19 mentions so as to robustly address that comment (see response above). 
The manuscript text mentioned in C17-C18 was just linking this paper to other prior work on ATAs. 
It's not necessary for the reader to understand that work to understand this paper, the point is just to make 
the link to that other literature, so that readers can explore it if they so choose, 
and to show that ATAs have been found to be important in ecology, generally, to motivate the study.
Furthermore, we don't think our paper is any more specialized or mathematical than the papers Ellner et al (2016, 2019) which
we cite, and which inspired our paper, and which were published in Ecology Letters. Nor is it more 
specialized than many other papers we have read in Ecology Letters. Quantitive approaches are absolutely 
central in ecology these days, and increasingly so. So comments which would tend to sideline such paper
"specialist journals", which fewer people read, seem outdated to us, especially when the papers in question 
are about mainstream topics such
as coexistence. Nevertheless, we have gone through the paper and redoubled our efforts to make the 
paper broadly accessible.} \\
%We have to go through the paper and redoubled our efforts to make the paper broadly accessible.

@@

\noindent C22) L74: The optimum might actually be the same, but they need different responses. \\

\noindent ***\emph{Thanks for pointing this out, we have corrected our language.} \\
%DONE

\noindent C23) Eq. 1: I would not use any equations in the introduction, and in general this part of the introduction did not feel like an introduction at all and might require a different sub-section \\

\noindent ***\emph{We disagree, and would prefer not to make this change, for several reasons. 
First, the lottery model is a very simple model, at the same level of complexity
as models introduced in essentially all introductory 
undergraduate ecology courses (e.g., the Lotka-Volterra competition model), so there is no argument that 
the readership of Ecology Letters cannot easily understand the material presented here.
Second, this material serves a clear motivating purpose, most appropriately placed in the Introduction. 
The last author (DCR) has found, from decades of pitching 
quantitative papers to general ecological audiences, that clear motivation is essential - it is crucial to demonstrate 
clearly that an effect one plans to elaborate is present, and potentially important, prior to asking the reader to engage 
the kind of mental energy necessary to read the guts of the paper. So this 
paragraph is for showing that the effect of ATAs on coexistence can be real, at least for a simple 
example. That goal is convincingly 
accomplished by the paragraph, more so than could be accomplished by a purely qualitative, words-only Introduction. 
Given that the math is simple, and also more precise and definitive than 
qualitative verbal arguments, we think this material is actually
more beneficial to the readership and more effective for the intended motivational purpose than would be more words
(and we see based on C25 below that the referee appears to agree with us here in some respects, at least). 
Third, quantitative approaches are central in ecology these days, and increasingly so. The automatic tendency
to want to exclude mathematics of any kind from the Introduction, especially when the math is simple,
accessible, and serves a clear motivational purpose, seems like a dated bias to us. 
Finally, we firmly believe that ecology as a discipline gets stronger as it continues the clear
trend of the last few decades of incorporating more quantitative approaches. We believe we
benefit the field by challenging assumptions and biases such as categorically not putting equations 
in the Introduction.} \\
%Jasmin to read the above and let Dan know if it's convincing or not, and how it can be improved.
%It's done except for any improvements or revisions that Jasmin may suggest.

@@

\noindent C24) L93: I suggest ``Note that $N = N_1(t) + N_2(T)$ is constant through time''. \\

\noindent ***\emph{OK, we made this change, even though the same equation already appears just a few lines
above this spot. Perhaps the referee is right that the reminder will benefit some readers.} \\
%DONE

\noindent C25) L95-99: I think this is a great teaser to get the readers to later dive in into your mathematics \\

\noindent ***\emph{Thank you! This is exactly our purpose, here, and we are pleased it has been effective.
Relating back to our response to C23, we think this purpose cannot really be convincingly accomplished
without first stating the lottery model, since the result uses that model. Math is more precise and can be
more convincing than language, the trade off being that some readers won't understand some math. However, the 
math we have in the Intro is so simple we think all readers of Ecology Letters will understand it, so there are
benefits and basically no costs of including it there. } \\
%DONE

@@

\noindent C26) L126: Do you do this in the appendix? \\

\noindent ***\emph{No. We can see how it would be done, but we have not formalized the mathematics. 
To avoid making claims that we have not fully supported
with rigorous mathematics, we rephrased to say the iid assumption is 
``probably not necessary''. Thanks for spurring us to do this.} \\
%DONE

@@

\noindent C28) Eq. 6: Maybe highlight that you use different brackets ([EC] versus (EC)), or probably even better, make 5 and 6 one equation using curly under brackets. \\

\noindent ***\emph{Yes, it is a good idea to make it completely clear to the reader that these are 
used distinctly. We now do so, though we do it above this point in the text - it is better, in our view,
to do this the first time we use the hard brackets. So thanks for pointing this out.
We do not know what the referee means by ``using curly under brackets'' so have not done anything
in response to that part of this comment.} \\
%DONE

\noindent C29) L230 $\mu_1-\mu_2$ is negative, why not take $\mu_2-\mu_1$? \\

\noindent ***\emph{Well, we can do this if the editor insists, but to us, avoiding the use of
negative numbers is insufficient reason to have to undertake the substantial and time-consuming 
work (and potential for error) 
of re-doing several figures, and revising many formulas (in the main text and SI) which currently use the other convention.
After all, negative numbers are basic, and represent no impediment to the average ecologist understanding the 
paper.} \\
%DONE

@@

\emph{Regarding the second part of the referee comment, pertaining to Figure 1d and e, it is perplexing to us that 
this is mentioned here,
as the paragraph at L316 is not directly related to Fig. 1d-e. That figure
shows motivational empirical examples, distinct from the lottery model work discussed around L316. Nevertheless, we take
seriously the referee's comment that they are are not convinced there are ATAs in Fig. 1d-e, and ...
%Jasmin, let's discuss what to do here. We could work back to what specific result this is in 
%the original paper and extract from p-value info, or we could come up with a better example,
%or something else, not sure, let's discuss. Maybe also augment the figure to make it clearer?
%How about if Jasmin: 1) adds statistical support to caption; 2) adds to the figure some means of
%hopefully drawing people's attention to the right features of this thing.
} \\

\noindent C31) Figure 2: I recommend to add a line at $y = 0$. Also the different subgraphs have different $y$-scales which I’m not sure is beneficial, but unsure myself. \\

\noindent ***\emph{There is already a line at $y=0$. You can see it, for instance, on panel (e). 
It gets unavoidably overplotted on panels for which one of the plotted quantities is close to 0. The reason
for different $y$-axis extents is so that you can see the variation within each panel with maximum 
clarity. The main point, here, is to compare the relative contributions of different components
to coexistence, so it's important to show within-plot variation (between lines and across 
values of the $x$-axis) clearly. We tried plotting with the same scales everywhere and it obscured some of the 
patterns. We thank the referee for expressing their own uncertainty, which was
useful to be aware of.} \\
%DONE

@@

\noindent C32) Figure 2 and 3: It appears that ATA is almost never the relevant factor, however I wonder whether this is because ATA is relatively weak, or whether the same analysis of say $\delta_i$ would lead to a similar pattern, it might be worth creating these graphs in the appendix to strengthen your case. \\

\noindent ***\emph{For Fig. 2, the referee is right that ATAs are only rarely important, as explained in the text 
(see first paragraph of section 4.1). However, for Fig 3, the referee is incorrect. In that figure, the ATA term
is usually one of the largest-magnitude contributions to coexistence. Our results as to the importance of ATAs for coexistence 
contrast between our log-normal and beta funcundities lottery models, as explained in the last paragraph
of section 4.1. That's one of the points we are making.} \\
%DONE

@@

\noindent C39) ...but how ATAs enter in play is not well solved. It is just mentioned that ATAs can modify the EC covariation, which I believe is not enough, a reader might not understand later the explanation when the lottery model is presented. \\

\noindent ***\emph{response} \\
%Same comment again, Jasmin can work on it.

\noindent C40) ATAs can not only modify the importance of the storage effect but also of other mechanisms such as relative non-linearity, which is based on how competition fluctuates in ways that favor the competitor. Could the authors briefly indicate how this could occur? \\

\noindent ***\emph{This cannot occur. ATAs can only influence storage effects, not the other mechanisms, as stated at
the top of section 2.4 and explained mathematically in the SI section cited there. Basically, as explained in 2.4, 
this is because ATAs are about covariation between $E$ and $C$, whereas the mechanisms other than storage effects
do not pertain at all to covariation. For instance, relative nonlinearity is about the effects of variation 
in $E$ or $C$, not covariation between them.} \\
%DONE

@@

\noindent C41) The lottery model is used to illustrate the importance of ATAs, and two types of fecundities are taken into account (log-normal and beta), can the author explain why both are important, and later on the results and discussion why ATAs play a major role with beta fecundities and not with log-normal \\

\noindent ***\emph{<This is a good question. We need to find a place where the models are presented to motivate
them, AND a place later, in the Discussion (or a new SI section called ``Additional discussion points'') where
we talk about why ATAs matter for one but not the other. I think this is just because the
effects of extremes are accentuated in the beta one, and so the effects of associations in the extremes are
accentuated. But that's not unreasonable, since many species have environments which fluctuate between
being suitable and unsuitable.>} \\
%Jasmin to consider this while brainstorming the prior comment. Dan can help, or do this one, but 
%I'll wait until after we've met and had a chance to discuss intutive explanations of ATA effects
%that Jasmin has been working on.

@@

\noindent C42) The model used for the chemostat is in its formulation quite different compared to the lottery model. Although the technical details are left for the supplementary material, can the authors explain more in the detail in the main text how the importance of ATAs can be used for this particular model. \\

%This referee is asking us to explain a lot of things in more detail. We have to be careful coming across as
%though we have written a paper ``not for a mainstearm audience'', but we can decline some of these requests
%at least for the main text) as not possible because of space constraints. This particular request does not 
%make perfect sense because of grammatic problems. But overall the referee seems to want us to explain the
%effect intuitively, whereas all we've really done is demonstrate its importance. If we want to insert an intuitive 
%explanation, that'd be more than we've tried for. We should talk about whether we have or could make
%something like that.
\noindent ***\emph{response} \\
%Jasmin to also lump this in with her thinking about the earlier ref 3 comments. 
%Q: When you look at E and C marginals for the diatom model, do they have most of their mass in the tails, like the 
%beta-fecundidies lottery model, or most of their mass in the middle, like the normal-fecundities lottery model?
%Let's talk more about this after Jasmin communicates to Dan the work she's done explaining the ATA effect
%intuitively.

### 2023-07-19 paper/Paper.Rnw
continued addressing referee comments, about done with what I can do before meeting with Jasmin

%Currently 150 of an allowed 150 words for Ecology Letters
\begin{abstract}
\noindent Species coexistence attracts wide 
interest in ecology. Modern coexistence
theory (MCT) identifies coexistence mechanisms,
one of which, storage effects, hinges on relationships between fluctuations
in environmental and competitive pressures. However, such relationships
are typically measured using covariance, which does not account for 
the possibility that environment and competition may be more related to each other
when they are strong than when weak, or vice versa. 
Recent work showed that such `asymmetric tail associations' (ATAs) are common
between ecological variables, and are important for 
extinction risk, ecosystem stability, and other phenomena. 
We extend the MCT, decomposing storage effects to show the influence of ATAs.
Analysis of a simple model
and an empirical example using diatoms
illustrate that ATA influences can be comparable in magnitude to other
mechanisms of coexistence, and that ATAs can make the 
difference between species coexistence and competitive exclusion. 
ATA influences are an important new mechanism of coexistence. 
\end{abstract}

@@

%Intro the paper by intro'ing ATAs
At the core of ecology is the study of relationships between organisms and their environments.
Ecologists may ask, for instance, how local environment relates to 
population growth rate. To study such relationships and to understand how two variables interact, ecologists 
very commonly use metrics such as correlation and covariance. But these standard measures of the strength 
and direction of association between variables often do not tell the whole story 
\citep{Ghosh_2020_AER}.
To illustrate, the three pairs of variables in Fig. \ref{fig:pedagog1}a-c all have the same 
Pearson correlation despite striking differences in association structure: Fig. \ref{fig:pedagog1}b shows a 
symmetric type of association, while Fig. \ref{fig:pedagog1}a,c exhibit what has previously been referred to as 
``asymmetric tail association'' [ATA; \citealp{Ghosh_2020_AER}]. 
Metrics of association other than correlations and covariance are also often used (e.g., various forms of 
regression), but correlations are extremely common, and other common methods also reveal only a small portion 
of the information content of the relationship between variables \citep{Nelsen2006}.
ATAs and related ideas had previously been seldom considered in ecology, but recently several 
insights have been gained by studying how features of associations beyond standard measures 
impact various ecological phenomena \citep{deValpine_2014, Anderson_2019, Popovic_2019, Ghosh_2020_AER, Ghosh_2020_Eco,
Ghosh_2020_ecolevol, Ghosh_2021}. For instance, \cite{Ghosh_2020_Eco} argue that extinction risks can become elevated 
when metapopulations are subject to ATAs; and \cite{Ghosh_2021} show that ATAs can alter the stability of 
ecosystem functioning. Fig. \ref{fig:pedagog1}d,e show contrasting examples of ATAs in nature using 
plankton population density time series; and many other examples have also been previously explored 
\citep{Ghosh_2020_AER,Ghosh_2020_ecolevol,Ghosh_2021}. To our knowledge, the implications of ATAs for 
competition and coexistence have not yet been investigated. Doing so is the overarching goal of this study.

@@

%Summary-type paragraph of main results
We produced and applied a new extension of modern coexistence theory (MCT) that 
quantifies the impact of asymmetric tail associations (ATAs) on species coexistence. 
Building upon a recent, simulation-based approach to MCT \citep{Ellner_2016, Ellner_2019},
we decomposed storage effects into two new mechanisms, the contribution of ATAs to coexistence,
and the contribution of $EC$ correlation \emph{per se}.
Applications of our extended theory to the lottery model and to a laboratory diatom system demonstrated that  
ATA contributions to storage effects and to species coexistence 
can often be comparable in magnitude to the contributions of previously recognized mechanisms.
Our results add
coexistence and biodiversity maintenance to the list of ecological theories and phenomena on which 
ATAs have been shown to have an important impact \citep{Ghosh_2020_AER, Ghosh_2020_Eco, Ghosh_2020_ecolevol, Ghosh_2021}.  
Though storage effects are typically described as the contribution
of $EC$ \emph{covariance} to coexistence \citep{Chesson_2012, Ellner_2016, Ellner_2019},
our results suggest that storage effects should be redefined to constitute the 
contribution of \emph{associations} between $E$ and $C$, generally, including both ATA effects and
correlation \emph{per se}. The prior terminology, \emph{covariance}, is typically construed 
as reflecting the linear relationship between variables, ignoring complexities of the association,
such as ATAs, which we now know can be important. ATAs contributions to coexistence were less important than other mechanisms 
in some contexts (the log-normal fecundities lottery model), but were very important in other contexts
(the beta fecundities lottery model), including our empirically
based example (the diatom system). Though future work should seek to understand precisely when ATAs are or are not important
for coexistence, our work demonstrates the overall importance of this new mechanism.

@@

%Differences with the original, analytic version of the MCT
Our approach to exploring the influence of ATAs on coexistence relies on the computational
extension of the MCT of \cite{Ellner_2019},
and our results help demonstrate strengths of that approach.
Storage effects as computed using the original, analytic approach to MCT 
\citep{Chesson_1994,Chesson_2000} should equal our term $\Delta^{[E \nat C]}$, 
which we refer to as the contributions to coexistence of correlation \emph{per se}.
This is because the weak-noise/Taylor approximation approach used in the original 
MCT neglects the effects of ATAs when higher-order terms
in Taylor expansions are dropped. The term $\Delta^{(E \shp C)}$, which our theory 
inherits from \cite{Ellner_2016}, is also dropped. These observations do not diminish 
the major strengths of the 
original, analytic approach, which are well known and also elaborated by \cite{Ellner_2019};
one such strength is the generality that analytic approaches provide.
It may be possible to expand the mathematics of the original MCT to consider ATAs.
Additional discussion points are in SI sections \ref{SI-sect:vital_rates}-\ref{SI-sect:threeway}.

>>>>

%Currently 150 of an allowed 150 words for Ecology Letters
\begin{abstract}
\noindent Species coexistence attracts wide 
interest in ecology. Modern coexistence
theory (MCT) identifies coexistence mechanisms,
one of which, storage effects, hinges on relationships between fluctuations
in environmental and competitive pressures. However, such relationships
are typically measured using covariance, which does not account for 
the possibility that environment and competition may be more related to each other
when they are strong than when weak, or vice versa. 
Recent work showed that such `asymmetric tail associations' (ATAs) are common
between ecological variables, and are important for 
extinction risk, ecosystem stability, and other phenomena. 
We extend the MCT, decomposing storage effects to show the influence of ATAs.
Analysis of a simple model
and an empirical example using diatoms
illustrate that ATA influences can be comparable in magnitude to other
mechanisms of coexistence, and that ATAs can make the 
difference between species coexistence and competitive exclusion. 
ATA influences may be an important new mechanism of coexistence. 
\end{abstract}

@@

%Intro the paper by intro'ing ATAs
At the core of ecology is the study of relationships between organisms and their environments.
Ecologists may ask, for instance, how local environment relates to 
population growth rate. To study such relationships and to understand how two variables interact, ecologists 
very commonly use metrics such as correlation and covariance. But these standard measures of the strength 
and direction of association between variables often do not tell the whole story 
\citep{Ghosh_2020_AER}.
To illustrate, the three pairs of variables in Fig. \ref{fig:pedagog1}a-c all have the same 
Pearson correlation despite striking differences in association structure: Fig. \ref{fig:pedagog1}b shows a 
symmetric type of association, while Fig. \ref{fig:pedagog1}a,c exhibit what has previously been referred to as 
``asymmetric tail association'' [ATA; \citealp{Ghosh_2020_AER}]. 
Metrics of association other than correlations and covariance are also often used (e.g., various forms of 
regression), but correlations are extremely common, and other common methods also reveal only a small portion 
of the information content of the relationship between variables \citep{Nelsen2006}.
ATAs and related ideas had previously been seldom considered in ecology, but recently several 
insights have been gained by studying how features of associations beyond standard measures 
impact various ecological phenomena \citep{deValpine_2014, Anderson_2019, Popovic_2019, Ghosh_2020_AER, Ghosh_2020_Eco,
Ghosh_2020_ecolevol, Ghosh_2021}. For instance, \cite{Ghosh_2020_Eco} argue that extinction risks can become elevated 
when metapopulations are subject to ATAs, e.g., if environmental ``catastrophes'' are more spatially synchronous 
across component populations
than are years of good environment, limiting the potential for rescue effects; 
and \cite{Ghosh_2021} show that ATAs can alter the stability of 
ecosystem functioning. Fig. \ref{fig:pedagog1}d,e show contrasting examples of ATAs in nature using 
plankton population density time series; and many other examples have also been previously explored 
\citep{Ghosh_2020_AER,Ghosh_2020_ecolevol,Ghosh_2021}. To our knowledge, the implications of ATAs for 
competition and coexistence have not yet been investigated. Doing so is the overarching goal of this study.

@@

%Summary-type paragraph of main results
We produced and applied a new extension of modern coexistence theory (MCT) that 
quantifies the impact of asymmetric tail associations (ATAs) on species coexistence. 
Building upon a recent, simulation-based approach to MCT \citep{Ellner_2016, Ellner_2019},
we decomposed storage effects into two new mechanisms, the contribution of ATAs to coexistence,
and the contribution of $EC$ correlation \emph{per se}.
Applications of our extended theory to the lottery model and to a laboratory diatom system demonstrated that  
ATA contributions to storage effects and to species coexistence 
can often be comparable in magnitude to the contributions of previously recognized mechanisms.
Our results add
coexistence and biodiversity maintenance to the list of ecological theories and phenomena on which 
ATAs have been shown to have an important impact \citep{Ghosh_2020_AER, Ghosh_2020_Eco, Ghosh_2020_ecolevol, Ghosh_2021}.  
Though storage effects are typically described as the contribution
of $EC$ \emph{covariance} to coexistence \citep{Chesson_2012, Ellner_2016, Ellner_2019},
our results suggest that storage effects should be redefined to constitute the 
contribution of \emph{associations} between $E$ and $C$, generally, including both ATA effects and
correlation \emph{per se}. The prior terminology, \emph{covariance}, is typically construed 
as reflecting the linear relationship between variables, ignoring complexities of the association,
such as ATAs, which we now know can be important. ATAs contributions to coexistence were less important than other mechanisms 
in some contexts (the log-normal fecundities lottery model), but were very important in other contexts
(the beta fecundities lottery model), including our empirically
based example (the diatom system). 
Future work should seek to understand precisely when ATAs are or are not important
for coexistence. Our work demonstrates the potential for ATAs to be an important mechanism of coexistence, generally.

@@

%Differences with the original, analytic version of the MCT
Our approach to exploring the influence of ATAs on coexistence relies on the computational
extension of the MCT of \cite{Ellner_2019},
and our results help demonstrate strengths of that approach.
Storage effects as computed using the original, analytic approach to MCT 
\citep{Chesson_1994,Chesson_2000} should equal our term $\Delta^{[E \nat C]}$, 
which we refer to as the contributions to coexistence of correlation \emph{per se}.
This is because the weak-noise/Taylor approximation approach used in the original 
MCT neglects the effects of ATAs when higher-order terms
in Taylor expansions are dropped. The term $\Delta^{(E \shp C)}$, which our theory 
inherits from \cite{Ellner_2016}, is also dropped. These observations do not diminish 
the major strengths of the 
original, analytic approach, which are well known and also elaborated by \cite{Ellner_2019};
one such strength is the generality that analytic approaches provide.
It may be possible to expand the mathematics of the original MCT to consider ATAs.
Additional discussion points are in SI sections \ref{SI-sect:vital_rates}-\ref{SI-sect:storage_importance}.

### 2023-07-19 paper/SupMat.Rnw
continued addressing referee comments, about done with what I can do before meeting with Jasmin

>>>>

\section{   Diverse viewpoints on the potential general importance of storage effects and ATAs for coexistence}\label{sect:storage_importance}

Some researchers have the opinion that storage effects are likely to be unimportant for 
coexistence, generally, in field systems, though evidence is still insufficient to know for sure,
and other researchers have the reverse opinion. Regardless of how that debate is eventually resolved, 
ATAs can, \emph{a priori}, be important for coexistence even if storage effects are not.

Opinions differ on the likely general importance of storage effects for coexistence, and 
evidence so far is insufficient to clearly determine their importance.  
\cite{Zapeda_2019} states that ``evidence for the relative importance of the different mechanisms 
that promote species coexistence is lacking,'' though they went on to address this lack by showing that 
niche differentiation and relative nonlinearity were more important for species coexistence 
in a grassland system than were storage effects.   
\cite{ChuAdler2015} argued that ``Niche differences and average fitness differences jointly determine 
coexistence. However, little empirical information about the magnitude of these two mechanisms is available.''
They then went on to show strong niche differences and modest average fitness differences in five 
North American grassland and shrubland plant communities. During this process they found little 
contribution of fluctuation-dependent stabilization mechanisms such as storage effects to coexistence,
but they also stated that ``The temptation is to conclude that fluctuation-dependent mechanisms 
are weak in these ecosystems (e.g., Adler et al. 2010). However, we now believe that the models 
we fit may not be appropriate for rigorously quantifying fluctuation-dependent mechanisms.'' 
In a very recent and thorough paper, \cite{StumpVasseur2023} state that they know of only four studies that
used empirically quantified models to measure the relative importance of storage effects versus other 
coexistence mechanisms \citep{Descamps_2005,Letten2018,Hallett_2019,Zapeda_2019}. \cite{StumpVasseur2023} 
state that ``In all of these studies, the stabilizing contribution of the storage effect was on par 
with the contribution of relative nonlinearity.'' \cite{StumpVasseur2023} interpret this to be
evidence that storage effects are typically unimportant, because relative nonlinearity has been 
thought to be weak -- and they spend most of their paper giving 
a theoretical/conceptual explanation for the unimportance of storage effects. 
But they also admit the possibility that relative nonlinearity is 
stronger than has previously been appreciated, and that both storage effects and relative nonlinearity
provide appreciable contributions to coexistence. \cite{StumpVasseur2023} also state that all studies
that have measured both storage effect and variation-independent mechanisms have found that the latter
overshadow the former, though there have only been two such studies \citep{Hallett_2019,Zapeda_2019}.

Regardless of the general importance of storage effects for coexistence,  
ATAs can, \emph{a priori}, be important for coexistence even if storage effects are not.  
Storage effects are ATA effects plus what we have called the effects of $EC$ correlation \emph{per se} 
($\Delta_i^{[E \nat C]}$ in our notation). Either of these two summands can be negative or positive.
So if the two effects are opposite in sign and similar in magnitude, storage effects will be unimportant for 
coexistence, but ATA effects will be important. 
This is precisely what we observed in some of our lottery model results (e.g., the beta-fecundities model 
with left-tail associations).
It seems to us that only further research can 
clearly resolve the question of the general
relative importance of storage effects, ATA effects, and other mechanisms of coexistence.
